Limited plasticity but increased variance in physiological rates across ectotherm populations under climate change

### 2024-12-18 ms.qmd
Added models for acute-acclim effect size for overall habitat differences and to test climate variability hypotheses

  lnRRQ10_data_acc_minus_aclim <- data_wide %>%        
                  dplyr::select(!all_of(drop)) %>%  
                  filter(!is.na(lnRR_Q10_acute_t1) | !is.na(lnRR_Q10_acute_t2)) %>% 
                  mutate(lnRR_acute = weighted_avg_effects(lnRR_Q10_acute_t1, lnRR_Q10_acute_t2, V_lnRR_Q10_acute_t1, V_lnRR_Q10_acute_t2),
                      lnRR_acute_SV = weighted_avg_sv(V_lnRR_Q10_acute_t1, V_lnRR_Q10_acute_t2))   %>% 
                  dplyr::select(-lnRR_Q10_acute_t1, -lnRR_Q10_acute_t2, -V_lnRR_Q10_acute_t1, -V_lnRR_Q10_acute_t2)  %>% 
                  mutate(lnRRQ10_ac_acclimate = lnRR_acute - lnRR_Q10_acclim, 
                         lnRRQ10_ac_acclimate_SV = lnRR_acute_SV + V_lnRR_Q10_acclim) %>%
                  dplyr::select(-lnRR_Q10_acclim, -V_lnRR_Q10_acclim, -lnRR_acute_SV, -lnRR_acute, -grep("r", colnames(.)))  %>% 
                  mutate(match = paste(gsub("_", "", species_full), lat, long, sep = "."), 
                    max_temp_c = scale(temp_2, scale = FALSE))

@@

>>>>

  lnRRQ10_data_acc_minus_aclim <- data_wide %>%        
                  dplyr::select(!all_of(drop)) %>%  
                  filter(!is.na(lnRR_Q10_acute_t1) | !is.na(lnRR_Q10_acute_t2)) %>% 
                  mutate(lnRR_acute = weighted_avg_effects(lnRR_Q10_acute_t1, lnRR_Q10_acute_t2, V_lnRR_Q10_acute_t1, V_lnRR_Q10_acute_t2),
                      lnRR_acute_SV = weighted_avg_sv(V_lnRR_Q10_acute_t1, V_lnRR_Q10_acute_t2))   %>% 
                  dplyr::select(-lnRR_Q10_acute_t1, -lnRR_Q10_acute_t2, -V_lnRR_Q10_acute_t1, -V_lnRR_Q10_acute_t2)  %>% 
                  mutate(lnRRQ10_ac_acclimate = lnRR_acute - lnRR_Q10_acclim, 
                         lnRRQ10_ac_acclimate_SV = lnRR_acute_SV + V_lnRR_Q10_acclim) %>%
                  dplyr::select(-lnRR_Q10_acclim, -V_lnRR_Q10_acclim, -lnRR_acute_SV, -lnRR_acute, -grep("r[12]", colnames(.)))  %>% 
                  mutate(match = paste(gsub("_", "", species_full), lat, long, sep = "."), 
                    max_temp_c = scale(temp_2, scale = FALSE),
                           spp = species_full)

@@

# Create the phylo correlation matrix
phylo_cor_wild_ac_acclim <- vcv(tree_lnrr_wild, cor = TRUE)
        A_wild_ac_acclim <- vcv(tree_lnrr_wild, cor = FALSE) # For bayes, need cov matrix

# Scale data
        lnRRQ10_data_wild_acute_vs_acclim <- lnRRQ10_data_wild_acute_vs_acclim %>% 
                                                        mutate(cv_c = scale(cv),
                                                          acf_all_c = scale(acf_all),
                                                          c_accl_time = acclim_period - mean(acclim_period, na.rm = TRUE))

# How much data? Quite a bit with both acute and acclimation, which is good.
        lnRRQ10_data_wild_acute_vs_acclim %>% group_by(habitat) %>% summarise(n = n(), stdy = length(unique(record_num)))

# Plot the effects with climate variation and predcitabiliy

  lnRRQ10_data_wild_acute_vs_acclim  %>% 
    ggplot(aes(x = acf_all, y = lnRRQ10_ac_acclimate, color = habitat)) + geom_point() + facet_wrap(~habitat) + theme_classic() + my_theme() + labs(x = "Autocorrelation", y = "Acute - Acclimation", color = "Habitat") + geom_smooth(method = "lm", se = FALSE) + geom_smooth(method = "lm", se = FALSE, aes(group = habitat), linetype = "dashed") + theme(legend.position = "none")

  lnRRQ10_data_wild_acute_vs_acclim  %>% 
    ggplot(aes(x = cv, y = lnRRQ10_ac_acclimate, color = habitat)) + geom_point() + facet_wrap(~habitat) + theme_classic() + my_theme() + labs(x = "Coefficient of Variation (CV)", y = "Acute - Acclimation", color = "Habitat") + geom_smooth(method = "lm", se = FALSE) + geom_smooth(method = "lm", se = FALSE, aes(group = habitat), linetype = "dashed") + theme(legend.position = "none")

# How much missing data in acclimation time?
  lnRRQ10_data_wild_acute_vs_acclim %>% filter(is.na(acclim_period)) %>% group_by(habitat) %>% summarise(n = n(), stdy = length(unique(record_num)))

####################################################
  # Model3_new: Model environmental variability by acute/acclim
####################################################
  rerun = FALSE

# Ok, now that we have it we may as well just rerun the model using the effect size Reviewer 1 is talking about to show that it is nearly the same (rememnber, we lose power here). Also remoember that positive effect sizes show acclimation because it implies the acute slope is larger than the acclimation slope. Sp, we expect freshwater and marine to have positive overall effect sizes.

if(rerun){
model3_new_overall  <- metafor::rma.mv(lnRRQ10_ac_acclimate ~ max_temp_c + c_accl_time + habitat, V = lnRRQ10_ac_acclimate_SV,
                                       random = list(~1 | record_num,
                                                     ~1 | species_full,
                                                     ~1 | trait,
                                                     ~1 + max_temp_c| spp,
                                                     ~1 | obs),
                              R = list(species_full = phylo_cor_wild_ac_acclim), dfs = "contain", 
                              test = "t", data = lnRRQ10_data_wild_acute_vs_acclim)
  save(model3_new_overall, file = "./output/models/lnRRQ10_3_new_overall")
} else{
  model3_new_overall <- readRDS("./output/models/lnRRQ10_3_new_overall")
}

if(rerun){
  # More realistic model with 
  model3_new  <- metafor::rma.mv(lnRRQ10_ac_acclimate ~ max_temp_c + c_accl_time + cv_c + habitat + habitat:cv_c, V = lnRRQ10_ac_acclimate_SV,
                                       random = list(~1 | record_num,
                                                     ~1 | species_full,
                                                     ~1 | trait,
                                                     ~1 + max_temp_c| spp,
                                                     ~1 | obs),
                              R = list(species_full = phylo_cor_wild_ac_acclim), dfs = "contain", 
                              test = "t", data = lnRRQ10_data_wild_acute_vs_acclim)

    model3_new_bayes <- brms::brm(lnRRQ10_ac_acclimate | se(sqrt(lnRRQ10_ac_acclimate_SV), sigma = TRUE) ~ max_temp_c + c_accl_time + cv_c + 
                                  habitat + habitat:cv_c + 
                                  (1 | record_num) + 
                                  (1 | gr(species_full, cov = A_wild_ac_acclim)) + 
                                  (1 + max_temp_c| spp) + 
                                  (1|trait), data2 = list(A_wild_ac_acclim = A_wild_ac_acclim), 
                                  family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), 
                                  control = list(adapt_delta = 0.98, max_treedepth = 12), 
                                  chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, data = lnRRQ10_data_wild_acute_vs_acclim)

  # Save model3_new
    saveRDS(model3_new, file = "./output/models/lnRRQ10_3_new")
    saveRDS(model3_new_bayes, file = "./output/models/lnRRQ10_3_new_bayes")

# Orchard plot model3_new
  orchaRd:orchard_plot(model3_new, )
